### 2022-09-12 dissertation/chapters/02-artificial-neural-networks.tex
📝 (review): First set of changes

\begin{quote}
    \textit{There are an estimated 100 billion neurons in the human brain. That
        means that there are approximately as many neurons in the human brain as
        there are stars in our galaxy. Every one of these neurons are highly
        interconnected with other neurons, yielding a synapse count of 100
        trillion. These neural structures are the result of millions of years
        of evolution. It is with these neural structures that we are able to think,
    learn and dream.}
\end{quote}

The human brain is by far one of the most complicated biological organs in
nature. The human brain is what gives us the ability to think and learn. The
field of \ac{ML} has taken a lot of inspiration from the biological brain which
lead to the development of \acp{ANN} \cite{ref:rosenblatt:1958}. \acp{ANN}
are used throughout this thesis as the model to be trained using \acp{HH}.
Understanding how \acp{ANN} work is an important component of this thesis.

@@

\begin{itemize}
    \item
    Section~\ref{sec:anns:bn} gives background information on the \ac{BN}.

    \item
    Section~\ref{sec:anns:an} introduces the \ac{AN} and the various
    components that it is made up of. Brief discussions follow on
    input, \index{weights}weights, \index{net input
    signal}net input signal, \index{biases}biases,
    \index{activation functions}activation functions, and
    output.

    \item
    Discussions on the \ac{BN} and the \ac{AN} set the tone for the \ac{ANN}
    which is discussed in detail in Section~\ref{sec:anns:ann}. Brief
    discussions follow on \ac{ANN} architecture, topology, and
    \acp{FFNN}.

    \item
    Section~\ref{sec:anns:training} gives details on the training process,
    training sets, \index{supervised learning}supervised learning,
    \index{error function}error functions, performance measures
    and stopping conditions.

    \item
    Finally, a brief summary of the chapter is given in
    Section~\ref{sec:anns:summary}.
\end{itemize}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{biological_neuron.eps}
    \caption[The \index{Biological Neuron}Biological Neuron]{An illustration of
    the \index{Biological Neuron}Biological Neuron.}
    \label{fig:biological_neuron}
\end{figure}

The main parts of the \ac{BN} to focus on includes the cell body, the
\index{dendrites}\textit{dendrites} and the \index{axon}\textit{axon} because these
components are directly modeled in the \ac{AN}. \Acp{NN} are formed by
neurons that connect to each other. This is known as
\index{synaptogenesis}\textit{synaptogenesis} \cite{ref:huttenlocher:1997}.
Connections are made between the \index{axons}axons and
\index{dendrites}dendrites of various neurons. Such a connection is referred to
as a \index{synapse}\textit{synapse}. It is through \index{synapses}synapses
that neurons communicate with each other. Communication takes place by
electro-chemical pulse and is often referred to as an \textit{activation} or
\textit{action potential}.  Communication signals propagate from the
\index{dendrites}dendrites, through the cell body to the axon of a neuron,
provoking a signal in the post-synaptic neuron \cite{ref:engelbrecht:2007}.
Neurons like these are massively interconnected.  The greater the connection
between two neurons, the stronger the communication.  Kennedy
\cite{ref:kennedy:2016} defines stronger \index{Synapse}synapses as ones that
contribute more depolarisation to the neural membrane upon activation than
weaker ones. Stronger \index{Synapse}synapses have a higher probability of
generating an action potential in the post-synaptic neuron. During activation,
the pre-synaptic neuron release neurotransmitters that bind to the post-synaptic
neuron \cite{ref:khanacademy:synapse}. The frequent release of these molecules
cause the \index{synapse}synapse to grow. Connections that grow over time yield
stronger signals, while connections that are weak propagate low intensity
signals and vanish over time. This is how the human brain learns and forgets.
Every time we reinforce a memory in our brains, we ``train'' the neuron and the
connection becomes stronger. This is known as \index{synaptic
plasticity}\textit{synaptic plasticity} \cite{ref:huttenlocher:1997}. this
reinforcement of connections plays a pivotal role in the way we algorithmically
train \acp{ANN}.

@@

The An \ac{AN} implements a nonlinear mapping from $\mathbb{R}^{I}$ to
$\mathbb{R}^{T}$, usually in the ranges $[0,1]$ or $[-1,1]$, depending on the
activation function used \cite{ref:engelbrecht:2007}. This mapping is given
below:

@@

\noindent
where $f_{AN}$ is the function mapping produced by the \ac{AN}, $I$
is the total number of dimensions of the input in real-number space $\mathbb{R}$
and $T$ is the total number of dimensions of the target (desired output) in
real-number space.

An illustration of the \ac{AN} is given below:

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{artificial_neuron.eps}
  \caption[The \index{Artificial Neuron}Artificial Neuron]{An illustration of
  the \index{Artificial Neuron}Artificial Neuron.}
  \label{fig:artificial_neuron}
\end{figure}

@@

    \item
    \index{net input signal}\textbf{Net input signal:} The \index{net input signal}netinput signl models
    the net resulting activations from all connected pre-synaptic neurons or
    environment sensors. The \index{net input signal}net input signal is
    represented by $net$.

@@

    \item
    \textbf{Output:} This output models the activation by the post-synaptic neuron. Output is
    represented by the output vector $\vec{y}$.
\end{itemize}

@@

Input signals are $I$-dimensional vectors of numerical values that are obtained
either through some environment sensor or from other \acp{AN}. Input signals
are often referred to as \index{features}\textit{features} or \index{independent
variables}\textit{independent variables} \cite{ref:francis:2001}. In this
thesis, a single input vector is referred to as a pattern. Input data must first
be pre-processed before it is presented to the \ac{AN}. Input pre-processing
techniques are discussed next.

@@

There are many techniques to consider when pre-processing input data. From
scaling to the correct ranges to dealing with incomplete, invalid or even
irrelevant data. For the purposes of this thesis, only data encoding and
normalisation is considered. These topics are now discussed in more detail.

@@

For qualitative data, such as class labels, the class labels have to be
converted to numerical representations. In general, class labels are encoded as
vectors \cite{ref:srinidhi:2018, ref:brownlee:2017:one-hot}. This thesis makes
use of \index{One-Hot Encoding}\textit{one-hot encoding} of class labels which
yields a $K$-dimensional output vector $\vec{y}$ of $0$'s and a single $1$. Each
element $y_k$ of $\vec{y}$ uniquely refers to a class. An activation, represented by a value
of $1$ at a given index thus represents a given class. Harris and Harris
\cite{ref:harris:2010} describe \index{One-Hot Encoding}one-hot encoding as a
group of vector elements where the representations of classes are represented
with a single high element (represented by $1$) and all the others low
(represented by $0$).

@@

Input data must be normalised and scaled appropriately. In general, input data
is normalised to a range that is appropriate for the \index{Activation
Function}activation function being used. This yields input data in a range that
is comparable to the output produced. For the purpose of this thesis, the
min-max scaler \cite{ref:al:2006} and standard score scaler \cite{ref:jain:2005}
are normalisation techniques that are considered. These are discussed next.

@@

The \index{standard score scaler}standard score scaler, also known as the
\index{z-score scaler}\textit{z-score scaler}, scales features by subtracting
the mean and scaling to the unit variance of each dimension $i$ for all input
patterns $\vec{x}_p, p \in \{1,2, \dots, P\}$. The \index{standard score
scaler}standard score scaler is used as a pre-processing technique for target
values when the \index{hyperbolic tangent}\textit{hyperbolic tangent}
\index{activation function}activation function\index{activation function} is
used. The normalised for of $x_{i,p}^{'}$ as determined by the \index{standard
score scaler}standard score scaler is given in below:

@@

Weight initialisation is the process by which candidate solutions (represented
by the weights of the \ac{AN}) to the problem is placed in the search-space.
Weight initialisation influences the speed of convergence, the probability of
convergence and the generalisation capabilities of \acp{ANN}
\cite{ref:fernandez:2001}. Finding the optimal initialisation values for weights
is non-trivial and can be seen as another optimisation problem. Extensive
research on this topic has been done in the past \cite{ref:de:2016,
ref:erdogmus:2003, ref:yam:2000}.

\index{Weight Initialisation}Weight initialisation is dependent on the
\index{activation function}activation function used. If weights are initialised
to values that are too small, the vanishing gradients problem can occur
\cite{ref:hanin:2018}. If weights are initialised to values that are too big,
output of the activation function would not be in the active range. This leads
to hidden unit saturation and exploding gradients can occur
\cite{ref:hanin:2018, ref:yadav:2018}.

Research have presented many initialisation techniques \cite{ref:erdogmus:2003}.
Some techniques have been developed to address the problems that are mentioned
above \cite{ref:yadav:2018}. However, this thesis will only focus on
\index{random uniform sampling}random uniform sampling, \index{Glorot uniform
sampling}\textit{Glorot uniform (Xavier)} sampling and \index{Glorot normal
sampling}\textit{Glorot normal} sampling \cite{ref:glorot:2010}. These
techniques are briefly discussed next.

@@

\index{Glorot normal sampling}Glorot normal sampling initialises weights by
sampling from a truncated normal distribution centred on a mean of $0$ and with
$\sigma = \sqrt{\frac{2}{I + K}}$, where $\sigma$ is the standard deviation of
the distribution. $I$ and $K$ are respectfully the number of input and output
units in the weight vector. \index{Glorot normal sampling}Glorot normal sampling
have been shown to decrease training time \cite{ref:glorot:2010}.

@@

The most common way by which the \index{net input signal}net input signal is
calculated is by means of \acp{SU} \cite{ref:engelbrecht:2007}, which
computes the \index{net input signal}net input signal as the weighted sum of all
input signals as shown below:

@@

A bias/threshold term $\theta$ is introduced to help translate the output of the
\index{activation function}activation function \cite{ref:benitez:1997}. The
value of $\theta$ must be learned during the training process along with the
weights of the \ac{ANN}. In order to simplify equations, augmented vectors are
used in which input and hidden layers have an additional neuron/unit, called the
\textit{bias unit} \cite{ref:engelbrecht:2007}. Augumented vectors that include
the bias unit to the net input signal is presented next.

@@

\index{activation function}Activation functions are used to model the action
potential of the \index{AN}\ac{AN} \cite{ref:ziv:1994, ref:hodgkin:1952}, and is
a function over the augmented \index{net input signal}net input signal. It is
the mechanism that models whether an \index{AN}\ac{AN} fires or not. When the
output produced by the activation function surpasses some threshold value
$\theta$, we consider that neuron to have fired an output signal.
\index{activation function}Activation functions model \textit{phase shift}. In
the context of classification problems, these \index{activation
function}activation functions form decision boundaries between classes. In the
context of regression problems, these \index{activation function}activation
functions try to approximate some function that maps the input data to some
target data.

@@

\begin{equation}
    f_{AN}: \mathbb{R}^{I} \rightarrow [0,1]
    \label{eq:an_function_mapping_0_1}
\end{equation}

@@

\begin{equation}
    f_{AN}: \mathbb{R}^{I} \rightarrow [-1,1]
    \label{eq:an_function_mapping_minus_1_1}
\end{equation}

\noindent Research in the past have presented many \index{activation
function}activation functions \cite{ref:karlik:2011} of varying ranges. However,
this thesis focuses on the \ac{ReLU} \cite{ref:jarrett:2009, ref:nair:2010}, the
\ac{LReLU} \cite{ref:maas:2013}, the \index{sigmoid}sigmoid \cite{ref:lecun:1988} and the
\index{hyperbolic tangent}hyperbolic tangent \cite{ref:lin:2008} \index{activation
function}activation functions. These \index{activation function}activation
functions are presented next.

@@

\noindent An advantage of \ac{ReLU} is that it is not susceptible to the
\index{vanishing gradient problem}vanishing gradients problem \cite{ref:xu:2015,
ref:maksutov:2018} which occurs when gradients in the first layers of a
multi-layer \ac{ANN} approach zero, and have no effect in the training process.

@@

The \ac{LReLU} \index{activation function}activation function is a variant of
the \ac{ReLU} \index{activation function}activation function \cite{ref:xu:2015}.
Similar to \ac{ReLU} it is not susceptible to the \index{vanishing gradients
problem}vanishing gradients problem and it does not suffer from the \index{dying
ReLU problem}dying \ac{ReLU} problem \cite{ref:agarap:2018}, which occurs when
the neurons that use \ac{ReLU} \index{activation functions}activation functions
become \textit{inactive} and only output $0$ due to a negative net input signal.

The \ac{LReLU} \index{activation function}activation function is defines as:

@@

An illustration of the \ac{LReLU} with various values for $\alpha$ and $\theta =
0$ is given in Graph~\ref{grf:leaky_relu} below:

\begin{graph}[H]
    \centering
    \input{./plots/leaky_relu.tex}
    \caption[The \ac{LReLU} \index{activation function}activation function]{An illustration of the
    \acs{LReLU} \index{activation function}activation function.}
    \label{grf:leaky_relu}
\end{graph}

@@

\index{sigmoid}Sigmoid is a very popular \index{activation function}activation
function, because it introduces non-linearity into the model. The
\index{sigmoid}sigmoid \index{activation function}activation function is defined
as:

@@

An illustration of the \index{sigmoid}sigmoid \index{Activation
Function}activation function with various values for $\lambda$ and $\theta = 0$
is given below:

\begin{graph}[H]
    \centering
    \resizebox{0.8\textwidth}{!}{\input{./plots/sigmoid.tex}}
    \caption[The \index{sigmoid}sigmoid \index{activation function}activation
    function]{An illustration of the \index{sigmoid}sigmoid \index{Activation
    Function}activation function.}
    \label{grf:sigmoid}
\end{graph}

@@

\noindent The \index{hyperbolic tangent}hyperbolic tangent \index{activation
function}activation function with various values for $\lambda$ and $\theta = 0$
is given below:

\begin{graph}[H]
    \centering
    \resizebox{0.8\textwidth}{!}{\input{./plots/tanh.tex}}
    \vspace{10pt}
    \caption[The \index{hyperbolic tangent}hyperbolic tangent \index{Activation
        Function}activation function]{An illustration of the \index{Hyperbolic
    Tangent}hyperbolic tangent \index{activation function}activation function.}
    \label{grf:hyperbolic_tangent}
\end{graph}

@@

Output signals are $K$-dimensional vectors of numerical values that are produced
by the \index{activation function}activation function. Output signals are often
referred to as \textit{predictions}. If target pre-processing was applied, it is
necessary to post-process the output data. A brief discussion on post-processing
techniques follow.

@@

Output post-processing is necessary in order to convert data back to their
original ranges. The post-processing techniques that are applicable depend on
the type of problem (regression or classification), input pre-processing
techniques used, as well as the \index{activation function}activation function
used. For the purposes of this thesis, only data decoding, normalisation and
re-scaling techniques are considered.

@@

Normalisation of the output data is used to simplify the decoding process. This
thesis considers \index{softmax}\textit{softmax} and
\index{Argmax}\textit{argmax} normalisation to assist the decoding process.

@@

\noindent Graph~\ref{grf:softmax_argmax} illustrates the comparison of
transformations of the output vector $\vec{y}$, where the
\index{sigmoid}sigmoid, the \index{softmax}softmax and the \index{argmax}argmax
\index{activation function}activation functions are used.

\begin{graph}[H]
    \centering
    \resizebox{0.8\textwidth}{!}{\input{./plots/output_modifiers.tex}}
    \caption[The results of \index{softmax}softmax and \index{Argmax}argmax]{An
        illustration of \index{softmax}softmax and \index{Argmax}argmax applied
    after the \index{sigmoid}sigmoid \index{activation function}activation function.}
    \label{grf:softmax_argmax}
\end{graph}

@@

\acp{ANN} are modelled after our biological brains. The human brain is
biologically and chemically incredibly sophisticated. Trying to model an entire
brain is at this stage not computationally feasible. Sandberg
\cite{ref:sandberg:2008} approximates a computational requirement of 256 000
terabytes/s to emulate the entire brain. Given our shortage in hardware
capabilities, \acp{ANN} have been successfully applied to a range of problem
classes. Engelbrecht \cite{ref:engelbrecht:2007} summarises some common problems
that are solved using \acp{ANN}. These include:

@@

\acp{ANN} are organised in layers where a single layer can contain multiple
\acp{AN}. Generally, each layer makes use of the same \index{activation
function}activation function, but this is not a requirement. Output from one
layer is propagated as input to the next layer. This thesis focusses on the
simplest architecture, containing three particular layers, including the input,
hidden and output layers. \acp{ANN} with this type of architecture is usually
referred to as \textit{shallow} \acp{NN}. The input, hidden and output layers
are discussed next.

The input layer contains the input data to the \ac{ANN}. Since the
input layer simply provides the the input data, some literature do not consider
the input layer as a layer \cite{ref:engelbrecht:2007}.

@@

There are many different topologies \cite{ref:miikkulainen:2010}, however, this
thesis only focusses on a \index{fully connected topology}fully connected
topology.

@@

\acp{FFNN} were the first and simplest \acp{ANN} developed
\cite{ref:schmidhuber:2015}. \acp{FFNN} incorporate the basic layers (input,
hidden and output layers) by arranging them in sequential order. In \acp{FFNN}
information moves forward, in one direction, from the input nodes, through the
hidden nodes and finally to the output nodes and depending on the optimisation
algorith used, error correction information can be backpropagated through the
network. It can be said that input is \textit{fed} forward, through the
\ac{ANN}. The \index{activation function}activation function used is problem
dependent. \acp{FFNN} incorporate fully connected topologies amongst the layers
such that no cycles form \cite{ref:zell:1994}.

An illustration of a \ac{FFNN} using a three-layered architecture and a fully
connected topology is given below:

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{feedforward_neural_network.eps}
    \caption[A \index{Feedforward Neural Network}Feedforward neural network]{An
    illustration of a \index{Feedforward Neural Network}feedforward neural network.}
    \label{fig:ffnn}
\end{figure}

>>>>

\begin{quote}
    \textit{There are an estimated 100 billion neurons in the human brain. That
        means that there are approximately as many neurons in the human brain as
        there are stars in our galaxy. Every one of these neurons is highly
        interconnected with other neurons, yielding a synapse count of 100
        trillion. These neural structures are the result of millions of years
        of evolution. It is with these neural structures that we are able to think,
    learn and dream.}
\end{quote}

The human brain is by far one of the most complicated biological organs in
nature. The human brain is what gives us the ability to think and learn. The
field of \ac{ML} has taken a lot of inspiration from the biological brain which
lead to the development of \acp{ANN} \cite{ref:rosenblatt:1958}. \acp{ANN}
are used throughout this dissertation as the model to be trained using \acp{HH}.
Understanding how \acp{ANN} work is an important component of this dissertation.

@@

\begin{itemize}
    \item
    \textbf{Section \ref{sec:anns:bn}} gives background information on the \ac{BN}.

    \item
    \textbf{Section \ref{sec:anns:an}} introduces the \ac{AN} and the various
    components that it is made up of. Brief discussions follow on
    input, \index{weights}weights, \index{net input
    signal}net input signal, \index{biases}biases,
    \index{activation functions}activation functions, and
    output.

    \item
    Discussions on the \ac{BN} and the \ac{AN} set the tone for the \ac{ANN}
    which is discussed in detail in \textbf{Section \ref{sec:anns:ann}}. Brief
    discussions follow on \ac{ANN} architecture, topology, and
    \acp{FFNN}.

    \item
    \textbf{Section \ref{sec:anns:training}} gives details on the training process,
    training sets, \index{supervised learning}supervised learning,
    \index{error function}error functions, performance measures
    and stopping conditions.

    \item
    Finally, a brief summary of the chapter is given in
    \textbf{Section \ref{sec:anns:summary}}.
\end{itemize}

@@

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.8\textwidth]{biological_neuron.eps}
    \caption[The \index{Biological Neuron}Biological Neuron]{An illustration of
    the \index{Biological Neuron}Biological Neuron.}
    \label{fig:biological_neuron}
\end{figure}

The main parts of the \ac{BN} to focus on include the cell body, the
\index{dendrites}\textit{dendrites} and the \index{axon}\textit{axon} because these
components are directly modeled in the \ac{AN}. \Acp{NN} are formed by
neurons that connect to each other. This is known as
\index{synaptogenesis}\textit{synaptogenesis} \cite{ref:huttenlocher:1997}.
Connections are made between the \index{axons}axons and
\index{dendrites}dendrites of various neurons. Such a connection is referred to
as a \index{synapse}\textit{synapse}. It is through \index{synapses}synapses
that neurons communicate with each other. Communication takes place by
electro-chemical pulse and is often referred to as an \textit{activation} or
\textit{action potential}.  Communication signals propagate from the
\index{dendrites}dendrites, through the cell body to the axon of a neuron,
provoking a signal in the post-synaptic neuron \cite{ref:engelbrecht:2007}.
Neurons like these are massively interconnected.  The greater the connection
between two neurons, the stronger the communication.  Kennedy
\cite{ref:kennedy:2016} defines stronger \index{Synapse}synapses as ones that
contribute more depolarisation to the neural membrane upon activation than
weaker ones. Stronger \index{Synapse}synapses have a higher probability of
generating an action potential in the post-synaptic neuron. During activation,
the pre-synaptic neuron release neurotransmitters that bind to the post-synaptic
neuron \cite{ref:khanacademy:synapse}. The frequent release of these molecules
cause the \index{synapse}synapse to grow. Connections that grow over time yield
stronger signals, while connections that are weak propagate low intensity
signals and vanish over time. This is how the human brain learns and forgets.
Every time we reinforce a memory in our brains, we ``train'' the neuron and the
connection becomes stronger. This is known as \index{synaptic
plasticity}\textit{synaptic plasticity} \cite{ref:huttenlocher:1997}. this
reinforcement of connections plays a pivotal role in the way we algorithmically
train \acp{ANN}.

@@

The \ac{AN} implements a nonlinear mapping from $\mathbb{R}^{I}$ to
$\mathbb{R}^{T}$, usually in the ranges $[0,1]$ or $[-1,1]$, depending on the
activation function used \cite{ref:engelbrecht:2007}. This mapping is given
below:

@@

\noindent
where $f_{AN}$ is the function mapping produced by the \ac{AN}, $I$
is the total number of dimensions of the input in real-number space $\mathbb{R}$,
and $T$ is the total number of dimensions of the target (desired output) in
real-number space.

An illustration of the \ac{AN} is given is Figure \ref{fig:artificial_neuron}.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=\textwidth]{artificial_neuron.eps}
  \caption[The \index{Artificial Neuron}Artificial Neuron]{An illustration of
  the \index{Artificial Neuron}Artificial Neuron.}
  \label{fig:artificial_neuron}
\end{figure}

@@

    \item
    \index{net input signal}\textbf{Net input signal:} The \index{net input signal}net input signal models
    the net resulting activations from all connected pre-synaptic neurons or
    environment sensors. The \index{net input signal}net input signal is
    represented by $net$.

@@

    \item
    \textbf{Output:} The output models the activation by the post-synaptic neuron. Output is
    represented by the output vector $\vec{y}$.
\end{itemize}

@@

Input signals are $I$-dimensional vectors of numerical values that are obtained
either through some environment sensor or from other \acp{AN}. Input signals
are often referred to as \index{features}\textit{features} or \index{independent
variables}\textit{independent variables} \cite{ref:francis:2001}. In this
dissertation, a single input vector is referred to as a pattern. Input data must first
be pre-processed before it is presented to the \ac{AN}. Input pre-processing
techniques are discussed next.

@@

There are many techniques to consider when pre-processing input data. From
scaling to the correct ranges to dealing with incomplete, invalid or even
irrelevant data. For the purposes of this dissertation, only data encoding and
normalisation is considered. These topics are now discussed in more detail.

@@

For qualitative data, such as class labels, the class labels have to be
converted to numerical representations. In general, class labels are encoded as
vectors \cite{ref:srinidhi:2018, ref:brownlee:2017:one-hot}. This dissertation makes
use of \index{One-Hot Encoding}\textit{one-hot encoding} of class labels which
yields a $K$-dimensional output vector $\vec{y}$ of $0$'s and a single $1$. Each
element $y_k$ of $\vec{y}$ uniquely refers to a class. An activation, represented by a value
of $1$ at a given index thus represents a given class. Harris and Harris
\cite{ref:harris:2010} describe \index{One-Hot Encoding}one-hot encoding as a
group of vector elements where the classes are represented
with a single high element (represented by $1$) and all the others low
(represented by $0$).

@@

Input data must be normalised and scaled appropriately. In general, input data
is normalised to a range that is appropriate for the \index{Activation
Function}activation function being used. This yields input data in a range that
is comparable to the output produced. For the purpose of this dissertation, the
min-max scaler \cite{ref:al:2006} and standard score scaler \cite{ref:jain:2005}
are normalisation techniques that are considered. These are discussed next.

@@

The \index{standard score scaler}standard score scaler, also known as the
\index{z-score scaler}\textit{z-score scaler}, scales features by subtracting
the mean and scaling to the unit variance of each dimension $i$ for all input
patterns $\vec{x}_p, p \in \{1,2, \dots, P\}$. The \index{standard score
scaler}standard score scaler is used as a pre-processing technique for target
values when the \index{hyperbolic tangent}\textit{hyperbolic tangent}
\index{activation function}activation function\index{activation function} is
used. The normalised form of $x_{i,p}^{'}$ as determined by the \index{standard
score scaler}standard score scaler is given in below:

@@

Weight initialisation is the process by which candidate solutions (represented
by the weights of the \ac{AN}) to the problem are placed in the search-space.
Weight initialisation influences the speed of convergence, the probability of
convergence and the generalisation capabilities of \acp{ANN}
\cite{ref:fernandez:2001}. Finding the optimal initialisation values for weights
is non-trivial and can be seen as another optimisation problem. Extensive
research on this topic has been done in the past \cite{ref:de:2016,
ref:erdogmus:2003, ref:yam:2000}.

\index{Weight Initialisation}Weight initialisation is dependent on the
\index{activation function}activation function used. If weights are initialised
to values that are too small, the vanishing gradients problem can occur
\cite{ref:hanin:2018}. If weights are initialised to values that are too big,
output of the activation function would not be in the active range. This leads
to unit saturation, and exploding gradients can occur
\cite{ref:hanin:2018, ref:yadav:2018}.

Research has presented many initialisation techniques \cite{ref:erdogmus:2003}.
Some techniques have been developed to address the problems that are mentioned
above \cite{ref:yadav:2018}. However, this dissertation will only focus on
\index{random uniform sampling}random uniform sampling, \index{Glorot uniform
sampling}\textit{Glorot uniform (Xavier)} sampling and \index{Glorot normal
sampling}\textit{Glorot normal} sampling \cite{ref:glorot:2010}. These
techniques are briefly discussed next.

@@

\index{Glorot normal sampling}Glorot normal sampling initialises weights by
sampling from a truncated normal distribution centred on a mean of $0$ and with
$\sigma = \sqrt{\frac{2}{I + K}}$, where $\sigma$ is the standard deviation of
the distribution. $I$ and $K$ are respectfully the number of input and output
units in the weight vector. \index{Glorot normal sampling}Glorot normal sampling
has been shown to decrease training time \cite{ref:glorot:2010}.

@@

The most common way by which the \index{net input signal}net input signal is
calculated is by means of \acp{SU} \cite{ref:engelbrecht:2007}, which
compute the \index{net input signal}net input signal as the weighted sum of all
input signals as shown below:

@@

A bias/threshold term $\theta$ is introduced to help translate the output of the
\index{activation function}activation function \cite{ref:benitez:1997}. The
value of $\theta$ must be learned during the training process along with the
weights of the \ac{ANN}. In order to simplify equations, augmented vectors are
used in which input and hidden layers have an additional neuron/unit, called the
\textit{bias unit} \cite{ref:engelbrecht:2007}. Augumented vectors that include
the bias unit to the net input signal are presented next.

@@

An \index{activation function}activation function is used to model the action
potential of the \index{AN}\ac{AN} \cite{ref:ziv:1994, ref:hodgkin:1952}, and is
a function over the augmented \index{net input signal}net input signal. It is
the mechanism that models whether an \index{AN}\ac{AN} fires or not. When the
output produced by the activation function surpasses some threshold value
$\theta$, we consider that neuron to have fired an output signal.
\index{activation function}Activation functions model \textit{phase shift}. In
the context of classification problems, these \index{activation
function}activation functions form decision boundaries between classes. In the
context of regression problems, these \index{activation function}activation
functions try to approximate some function that maps the input data to some
target data.

@@

\begin{equation}
    f_{AN}: \mathbb{R}^{I} \rightarrow [0,1],
    \label{eq:an_function_mapping_0_1}
\end{equation}

@@

\begin{equation}
    f_{AN}: \mathbb{R}^{I} \rightarrow [-1,1].
    \label{eq:an_function_mapping_minus_1_1}
\end{equation}

\noindent Research in the past has presented many \index{activation
function}activation functions \cite{ref:karlik:2011} of varying ranges. However,
this dissertation focuses on the \ac{ReLU} \cite{ref:jarrett:2009, ref:nair:2010}, the
\ac{LReLU} \cite{ref:maas:2013}, the \index{sigmoid}sigmoid \cite{ref:lecun:1988} and the
\index{hyperbolic tangent}hyperbolic tangent \cite{ref:lin:2008} \index{activation
function}activation functions. These \index{activation function}activation
functions are presented next.

@@

\noindent An advantage of \ac{ReLU} is that it is not susceptible to the
\index{vanishing gradient problem}vanishing gradients problem \cite{ref:xu:2015,
ref:maksutov:2018} which occurs when gradients in the first layers of a
multi-layer \ac{ANN} approach zero, and have no effect on the training process.

@@

The \ac{LReLU} \index{activation function}activation function is a variant of
the \ac{ReLU} \index{activation function}activation function \cite{ref:xu:2015}.
Similar to \ac{ReLU}, it is not susceptible to the \index{vanishing gradients
problem}vanishing gradients problem, and it does not suffer from the \index{dying
ReLU problem}dying \ac{ReLU} problem \cite{ref:agarap:2018}, which occurs when
the neurons that use \ac{ReLU} \index{activation functions}activation functions
become \textit{inactive} and only output $0$ due to a negative net input signal.

The \ac{LReLU} \index{activation function}activation function is defined as:

@@

An illustration of the \ac{LReLU} with various values for $\alpha$ and $\theta =
0$ is given in Figure \ref{fig:anns:activation_functions:leaky_relu}.

\begin{figure}[htpb]
    \centering
    \input{./plots/leaky_relu.tex}
    \caption[The \ac{LReLU} \index{activation function}activation function]{An illustration of the
    \acs{LReLU} \index{activation function}activation function.}
    \label{fig:anns:activation_functions:leaky_relu}
\end{figure}

@@

\index{sigmoid}Sigmoid is a very popular \index{activation function}activation
function, because it is a continuous differentiable approximation of the step function, which was used in the original perceptron model developed by \citeauthor{ref:rosenblatt:1957} \cite{ref:rosenblatt:1957}. The
\index{sigmoid}sigmoid \index{activation function}activation function is defined
as:

@@

An illustration of the \index{sigmoid}sigmoid \index{Activation
Function}activation function with various values for $\lambda$ and $\theta = 0$
is given in Figure \ref{fig:anns:activation_functions:sigmoid}.

\begin{figure}[htpb]
    \centering
    \resizebox{0.8\textwidth}{!}{\input{./plots/sigmoid.tex}}
    \caption[The \index{sigmoid}sigmoid \index{activation function}activation
    function]{An illustration of the \index{sigmoid}sigmoid \index{Activation
    Function}activation function.}
    \label{fig:anns:activation_functions:sigmoid}
\end{figure}

@@

\noindent The \index{hyperbolic tangent}hyperbolic tangent \index{activation
function}activation function with various values for $\lambda$ and $\theta = 0$
is given in Figure \ref{fig:anns:activation_functions:hyperbolic_tangent}.

\begin{figure}[htpb]
    \centering
    \resizebox{0.8\textwidth}{!}{\input{./plots/tanh.tex}}
    \vspace{10pt}
    \caption[The \index{hyperbolic tangent}hyperbolic tangent \index{Activation
        Function}activation function]{An illustration of the \index{Hyperbolic
    Tangent}hyperbolic tangent \index{activation function}activation function.}
    \label{fig:anns:activation_functions:hyperbolic_tangent}
\end{figure}

@@

Output signals are $K$-dimensional vectors of numerical values that are produced
by the \index{activation function}activation function. Output signals are often
referred to as \textit{predictions}. If target pre-processing was applied, it is
necessary to post-process the output data. A brief discussion on post-processing
techniques follows.

@@

Output post-processing is necessary in order to convert data back to their
original ranges. The post-processing techniques that are applicable depend on
the type of problem (regression or classification), input pre-processing
techniques used, as well as the \index{activation function}activation function
used. For the purposes of this dissertation, only data decoding, normalisation and
re-scaling techniques are considered.

@@

Normalisation of the output data is used to simplify the decoding process. This
dissertation considers \index{softmax}\textit{softmax} and
\index{Argmax}\textit{argmax} normalisation to assist the decoding process.

@@

\noindent Figure \ref{fig:anns:activation_functions:softmax_argmax} illustrates the comparison of
transformations of the output vector $\vec{y}$, where the
\index{sigmoid}sigmoid, the \index{softmax}softmax and the \index{argmax}argmax
\index{activation function}activation functions are used.

\begin{figure}[htpb]
    \centering
    \resizebox{0.8\textwidth}{!}{\input{./plots/output_modifiers.tex}}
    \caption[The results of \index{softmax}softmax and \index{Argmax}argmax]{An
        illustration of \index{softmax}softmax and \index{Argmax}argmax applied
    after the \index{sigmoid}sigmoid \index{activation function}activation function.}
    \label{fig:anns:activation_functions:softmax_argmax}
\end{figure}

@@

\acp{ANN} are modelled after our biological brains. The human brain is
biologically and chemically incredibly sophisticated. Trying to model an entire
brain is at this stage not computationally feasible. \citeauthor{ref:sandberg:2008}
\cite{ref:sandberg:2008} approximates a computational requirement of 256 000
terabytes/s to emulate the entire brain. Despite our shortage in hardware
capabilities, \acp{ANN} have been successfully applied to a range of problem
classes. \citeauthor{ref:engelbrecht:2007} \cite{ref:engelbrecht:2007} summarises some common problems
that are solved using \acp{ANN}. These include:

@@

\acp{ANN} are organised in layers where a single layer can contain multiple
\acp{AN}. Generally, each layer makes use of the same \index{activation
function}activation function, but this is not a requirement. Output from one
layer is propagated as input to the next layer. This dissertation focuses on the
simplest architecture, containing three particular layers, including the input,
hidden and output layers. \acp{ANN} with this type of architecture are usually
referred to as \textit{shallow} \acp{NN}. The input, hidden and output layers
are discussed next.

The input layer contains the input data to the \ac{ANN}. Since the
input layer simply provides the input data, some literature does not consider
the input layer as a layer \cite{ref:engelbrecht:2007}.

@@

There are many different topologies \cite{ref:miikkulainen:2010}, however, this
dissertation only focuses on a \index{fully connected topology}fully connected
topology.

@@

\acp{FFNN} were the first and simplest \acp{ANN} developed
\cite{ref:schmidhuber:2015}. \acp{FFNN} incorporate the basic layers (input,
hidden and output layers) by arranging them in sequential order. In \acp{FFNN}
information moves forward, in one direction, from the input nodes, through the
hidden nodes and finally to the output nodes, and, depending on the optimisation
algorith used, error correction information can be backpropagated through the
network. It can be said that input is \textit{fed} forward, through the
\ac{ANN}. The \index{activation function}activation function used is problem
dependent. \acp{FFNN} incorporate fully connected topologies amongst the layers
such that no cycles form \cite{ref:zell:1994}.

An illustration of a \ac{FFNN} using a three-layered architecture and a fully
connected topology is given in Figure \ref{fig:ffnn}.

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.8\textwidth]{feedforward_neural_network.eps}
    \caption[A \index{Feedforward Neural Network}Feedforward neural network]{An
    illustration of a \index{Feedforward Neural Network}feedforward neural network.}
    \label{fig:ffnn}
\end{figure}
